---
title: "Gapminder Exploratory Analysis"
author: Phil Chapman
date: 2019-01-28
output: 
    html_document:
        number_sections: yes
        theme: cosmo
        highlight: tango
        toc: yes
        toc_depth: 3
        code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This report contains a brief example analysis of the famous gapminder data. See [Managing Many Models](https://r4ds.had.co.nz/many-models.html) in the [R For Data Science book](https://r4ds.had.co.nz/) by [Hadley Wickham](https://twitter.com/hadleywickham) for further information on the data and techniques used in this analysis.

# Set up

## Load R libraries

```{r, message=FALSE}
library(gapminder)
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(broom)
```

# Explore data

## All data

Plot life expectancy over time for all countries on a single plot.  There is general upward trend it is difficult to discern how countries or continents might differ. 

```{r}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3) +
    theme_bw()
```

## Single country

Filter out data for a single country:

```{r}
nz_data <- gapminder %>%
  filter(country == 'New Zealand')
nz_data
```

Ploting the data for New Zealand shows that life expectancy has increased over time:

```{r}
ggplot(nz_data, aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3) +
    theme_bw()
```

## Fit a model

Fitting a linear model for the NZ data lets us to estimate the rate of increase of life expectancy:

```{r}
nz_model <- lm(lifeExp ~ year, data = nz_data)
nz_model
```

Extract coefficients:

```{r}
coef(nz_model)
broom::tidy(nz_model)
```

## Fit many models

To fit a linear model for each country:

```{r}
many_models <- gapminder %>%
  group_by(continent, country) %>%
  tidyr::nest() %>%
  mutate(mod = purrr::map(data, ~ lm(lifeExp ~ year, data = .)))
many_models
```

Look at the first country:

```{r}
many_models$data[[1]]
many_models$mod[[1]]

```

## Extract coefficients for many countries

Next we wish to extract the intercept and slope for each linear model, this will enable us to compare rate of increase of life expectancy across countries:

```{r}
all_countries_coefficients <- many_models %>%
  mutate(intercept = purrr::map_dbl(mod, ~ coef(.)['(Intercept)']),
         slope = purrr::map_dbl(mod, ~ coef(.)['year'])) %>%
  dplyr::select(-data, -mod) 
all_countries_coefficients
```

## Plot info on many models

Finally plot the rate of increase in life expectancy per country and colour by continent.  This shows that countries in Europe have shown a smaller increase in life expectancy than those in Asia, whereas countries in Africa have a more variable change.

```{r}
ggplot(all_countries_coefficients, aes(continent, slope, color = continent)) +
  geom_violin() +
  geom_point(position = position_jitter(width=0.3)) +
  theme_bw()
```


# Conclusion

In this brief analysis we have explored the gapminder data using the list-cols approach to fit models for different countries.  Plotting the rate of increase in life expectancy shows differences between continents: asia sees the biggest increases with europe the smallest, and with africa having a range of outcomes.

# Session Info

```{r}
sessionInfo()
```
